---
title: "Mouse Ileum Analysis"
author: "Zhaorong Li"
date: "Jan/11/2022"
output: 
  
  html_document:
    toc: true
    toc_float:
      toc_collapsed: true
      toc_depth: 4
    number_sections: true
    fig_caption: yes
    theme: cerulean
editor_options: 
  chunk_output_type: console
params:
  TF: ''
---


```{r setup, include=FALSE}
library(Seurat)
library(tidyr)
library(reshape2)
library(reactable)
library(RColorBrewer)
library(psych)
library(gridExtra)
```

# MERFISH Image visualization

This dataset [1] is downloaded from https://zenodo.org/record/5512749#.Yd3H0dHMKUl. The Baysor Cell segmentation and the author generated cell type assignment are used in the analysis.



## Plot Cell by centroid Coordinates
```{r Plot Cell by centroid Coordinates, fig.width=10,fig.height=10}

Spacial.information=read.csv('segmentation.csv.gz',header = T)

Spacial.information=Spacial.information[Spacial.information$is_noise!='true',]
Spacial.information$cell=as.character(Spacial.information$cell)
Cell_spacial_information=read.csv('segmentation_cell_stats.csv.gz',header = T,row.names = 1)

Celltype=read.csv('cell_assignment.csv.gz',header = T,row.names = 1)
colnames(Celltype)=c('Celltype')

Cell_spacial_information=merge(Cell_spacial_information,Celltype,by = 0)
colnames(Cell_spacial_information)=paste0('cell_',colnames(Cell_spacial_information))

colnames(Cell_spacial_information)[1]='cell'

Cell_spacial_information=Cell_spacial_information[Cell_spacial_information$cell_Celltype!='Removed',]
getPalette = colorRampPalette(brewer.pal(9, "Dark2"))


color=getPalette(length(unique(Cell_spacial_information$cell_Celltype)))

ggplot(Cell_spacial_information,aes(x=cell_x,y=cell_y,colour=cell_Celltype))+geom_point()+
  xlim(min(Cell_spacial_information$cell_x),max(Cell_spacial_information$cell_x))+
  ylim(min(Cell_spacial_information$cell_y),max(Cell_spacial_information$cell_y)) + scale_color_manual(values = color)+ theme_bw() + 
  theme(
    # get rid of panel grids
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    # Change plot and panel background
    plot.background=element_rect(fill = "black"),
    panel.background = element_rect(fill = 'black'),
    # Change legend 
    #legend.position = c(0.6, 0.07),
    #legend.direction = "horizontal",
    legend.background = element_rect(fill = "black", color = NA),
    legend.key = element_rect(color = "gray", fill = "black"),
    legend.title = element_text(color = "white"),
    legend.text = element_text(color = "white")
  )

```

## Plot Cell by real probes.
```{r Plot Cell by real probes, fig.width=10,fig.height=10}

Spacial.information=Spacial.information[Spacial.information$cell%in%rownames(Cell_spacial_information),]

Spacial.information=merge(Spacial.information,Cell_spacial_information,by = 'cell')

colnames(Spacial.information)[grepl('Celltype',colnames(Spacial.information))]='Celltype'


ggplot(Spacial.information,aes(x=x,y=y,colour=Celltype))+geom_point(shape='.')+
  xlim(min(Spacial.information$x),max(Spacial.information$x))+
  ylim(min(Spacial.information$y),max(Spacial.information$y)) + scale_color_manual(values = color)+ theme_bw() + 
  theme(
    # get rid of panel grids
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    # Change plot and panel background
    plot.background=element_rect(fill = "black"),
    panel.background = element_rect(fill = 'black'),
    # Change legend 
    #legend.position = c(0.6, 0.07),
    #legend.direction = "horizontal",
    legend.background = element_rect(fill = "black", color = NA),
    legend.key = element_rect(color = "gray", fill = "black"),
    legend.title = element_text(color = "white"),
    legend.text = element_text(color = "white")
  )

```

# Analysis

By looking at the visualization results from Part 1, we can see that the cellular composition of different parts of the tissue are different.

Can we use the unsupervised clustering algorithm to find different parts of the tissue?

Time to find out.

## Area Calculator Function

This function will break the Image into multiple small fragments. and calculate the cell composition in each small fragments.

    Logic:
      The whole image is broken into small fragments.

      for each small fragment:
  
        for each cell has overlap with the small fragment:
      
          if the cell has more than x% (x determined by the user) of cell's probes'areas in the small fragment, the cell is determined to be in the small fragment.
        Cell composition of the small fragment is determined by the cells assigned to it.
      
The percentage threshold and the size of small fragment can be determined by the user.

One tip: if the number of fragments is increased, please decrease the proportion_threshold. Otherwise many area will be deemed as no cells.


```{r Area Cellular composition Calculator}
source('functions.R')

```

```{r Area determination and Calculation}
Area.50.50=Area_Calculator(min(Spacial.information$x),
                                 max(Spacial.information$x),
                                 min(Spacial.information$y),
                                 max(Spacial.information$y),
                                 50,Spacial.information,proportion_threshold = 0.5)





reactable(Area.50.50)


```

## Normalization and Scaling

The cell composition of each fragment is scaled by the total number of cells in the fragment. The scaled number is multiplied by 1,000,000 and then log transformed.

Basically it is log transformed CPM normalized. This step is included in the Area_Cluster() function.

```{r Normalization and Scaling}

Area.50.50.normalized=Area.50.50/rowSums(Area.50.50)
Area.50.50.normalized=Area.50.50.normalized*(1e6)
Area.50.50.normalized=log(Area.50.50.normalized+1)

reactable(Area.50.50)

```

## Clustering

In this case I used the Louvain Community Clustering Algorithm from igraph package to cluster the area.


```{r clustering}

clusterlouvain <- Area_Cluster(Area.50.50,ScaleFactor = 1e6,0.5)

```

## Visualization of Clustering Results

```{r Visualization of Clustering Results, fig.width=20,fig.height=20}

Area.Cluster.assignment=clusterlouvain$Area.Cluster.assignment

Clustering.Distribution=clusterlouvain$Clustering.Distribution

melt=Area.Cluster.assignment
melt$cluster=as.factor(melt$cluster)
melt=melt(melt,id.vars='cluster')
colnames(melt)=c('Cluster','CellType','Cells')
Cluster_cell_composition=ggplot(melt,aes(x=Cells,y=Cluster,fill=CellType))+geom_bar(stat = 'identity',position = 'fill')+ scale_fill_manual(values = color)+ theme_bw()

centroid.graph=ggplot(Cell_spacial_information,aes(x=cell_x,y=cell_y,colour=cell_Celltype))+geom_point()+
  xlim(min(Cell_spacial_information$cell_x),max(Cell_spacial_information$cell_x))+
  ylim(min(Cell_spacial_information$cell_y),max(Cell_spacial_information$cell_y)) + scale_color_manual(values = color)+ theme_bw() +xlab('x') +ylab('y') +
  theme(legend.position = "none",
    # get rid of panel grids
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    # Change plot and panel background
    # Change legend 
    #legend.position = c(0.6, 0.07),
    #legend.direction = "horizontal",
)
Clustering.Distribution.plot=clusterlouvain$Clustering.Distribution.plot

grid.arrange(Clustering.Distribution.plot, centroid.graph,Cluster_cell_composition, ncol=2)

```

## Explore the Transcriptome Based on the Area Clustering Results

The Clustering algorithm separate the image into 5 clusters.

Cluster 1 and 2 are the center/inside of the tissue, which is rich in immune cells

Cluster 4 is the muscle lining of the Ileum.

Cluster 3 and 5 are the outside lining of the Villi. Cluster 5 is the tip of the villus which is rich in Top Villus Enterocytes. while Cluster 3 is rich in Mid Villus and Bottom Villus Enterocytes.


So in this case I would love to compare the difference between the gene expression profile of Enterocytes (Mid Villus) in Area clusters 4 and the gene expression profile of Enterocytes (Mid Villus) in other Area clusters.

```{r Assign Cells to Area Clusters}


Cell_spacial_information$Area_Cluster=''

for (i in 1:nrow(Cell_spacial_information)) {
  start=1
  stop=0
  while(start<=nrow(Clustering.Distribution)&stop==0) {
    if (Cell_spacial_information$cell_x[i]>Clustering.Distribution$xmin[start]&
        Cell_spacial_information$cell_x[i]<Clustering.Distribution$xmax[start]&
        Cell_spacial_information$cell_y[i]>Clustering.Distribution$ymin[start]&
        Cell_spacial_information$cell_y[i]<Clustering.Distribution$ymax[start]) {
      Cell_spacial_information$Area_Cluster[i]=Clustering.Distribution$cluster[start]
      stop=1
    }
    else {
      start=start+1
    }
  }
  rm(i)
}
write.table(Cell_spacial_information,'Cell_spacial_information.tsv',quote = F,sep = '\t',row.names = F)

reactable(Cell_spacial_information)

metadata=Cell_spacial_information
# Remove the cells that are not in the any Area cluster.
metadata=metadata[metadata$Area_Cluster!='',]

rownames(metadata)=metadata$cell
Counts=read.table('MERFISH/data_release_baysor_merfish_gut/data_analysis/baysor_membrane_prior/segmentation/segmentation_counts.tsv',header = F,row.names = 1,check.names = F)
colnames(Counts)=seq(1,ncol(Counts))
Counts=Counts[,rownames(metadata)]
Seurat_Object=CreateSeuratObject(Counts,meta.data = metadata)



Enterocyte=subset(Seurat_Object,subset=cell_Celltype%in%c('Enterocyte (Mid Villus)','Enterocyte (Bottom Villus)',
                                                          'Enterocyte (Top Villus)'))
Enterocyte=NormalizeData(Enterocyte)
Idents(Enterocyte)='Area_Cluster'
reactable(data.frame(table(Enterocyte$Area_Cluster)))
Enterocyte.Area.markers=FindAllMarkers(Enterocyte,test.use = 'MAST',only.pos = T)

reactable(Enterocyte.Area.markers)

```

The results show that the Apob and Slc51a are significantly upregulated in the Enterocyte in the outer linings of the villi, while gene Stmn1 is significantly downregulated.

Gene Introduction:

https://www.genecards.org/cgi-bin/carddisp.pl?gene=APOB

https://www.genecards.org/cgi-bin/carddisp.pl?gene=SLC51A&keywords=Slc51a

https://www.genecards.org/cgi-bin/carddisp.pl?gene=STMN1&keywords=Stmn1





# Conclusion

Coming soon.



# Reference

[1] Moffitt, Jeffrey, Xu, Rosalind, Kharchenko, Peter, Petukhov, Viktor, Cadinu, Paolo, Soldatov, Ruslan, & Khodosevich, Konstantin. (2021). MERFISH measurements in the mouse ileum [Data set]. https://doi.org/10.5061/dryad.jm63xsjb2










